---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# econdataverse

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/econdataverse)](https://CRAN.R-project.org/package=econdataverse)
<!-- badges: end -->

The 'EconDataverse' is a universe of open-source packages to work  seamlessly with economic data. This package is designed to make it easy to  install and load multiple 'EconDataverse' packages in a single step. Learn more about the 'EconDataverse' at [econdataverse.org](https://www.econdataverse.org).

## Installation

You can install the development version of `econdataverse` from [GitHub](https://github.com/tidy-intelligence/r-econdataverse) with:

``` r
# install.packages("pak")
pak::pak("tidy-intelligence/r-econdataverse")
```

## Usage

`library(econdataverse)` will load the core EconDataverse packages:

- `econid`, for standardizing identifiers.
- `econtools`, for combining data sources.
- `imfweo`, for the International Monetary Fund World Economic Outlook (WEO).
- `oecdoda`, for the OECD Official Development Assistance (ODA).
- `owidapi`, for the Our World in Data API.
- `uisapi`, for the UNESCO Institute for Statistics API.
- `wbids`, for the World Bank International Debt Statistics (IDS).
- `wbwdi`, for the World Bank World Development Indicators (WDI).

```{r}
library(econdataverse)
```

And you can check that all EconDataverse packages are up-to-date with `econdataverse_update()`:

```{r}
econdataverse_update()
```

## Contributing

Contributions to `econdataverse` are welcome! If you'd like to contribute, please follow these steps:

1. **Create an issue**: Before making changes, create an issue describing the bug or feature you're addressing.
2. **Fork the repository**: After receiving supportive feedback from the package authors, fork the repository to your GitHub account.
3. **Create a branch**: Create a branch for your changes with a descriptive name.
4. **Make your changes**: Implement your bug fix or feature.
5. **Test your changes**: Run tests to ensure your changes don't break existing functionality.
6. **Submit a pull request**: Push your changes to your fork and submit a pull request to the main repository.
